---
title: "Simulating multiple populations at once"
author: "Jaideep Joshi"
date: "28 March 2022"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(fisheRy)
library(tidyverse)
source("../tests/ref/parameters.cod.R")
source("../tests/ref/simulator.7.R")
```

```{r}
re_simulate = T

# if one of the files doesnt exist, re-run the simulations
re_simulate = re_simulate | !file.exists("outputs/res_ibm_scan_lmin.Rdata")
```

We can use the fishery simulator to simulate multiple populations (i.e., populations with different control parameters) at once. Suppose we want to look at fishery properties as a function of the harvest proportion. 

### Preparation

```{r}
params_file = "../params/cod_params_dankel22.ini"
```

As usual, we begin with creating a prototype fish.

Then, we create a simulator with this prototype, which will simulate a given population under different conditions.

```{r}
# Create a prototype fish as usual
fish = new(Fish, params_file)
```


### Create a population to simulate

Then we create a population that will be reused by the simulator.

```{r}
pop = new(Population, fish)
pop$readParams(params_file, F)
pop$set_superFishSize(5e6)
pop$verbose = T
pop$noFishingEquilibriate(5.61)
```



### Create a simulator

```{r}
sim = new(Simulator, fish)

sim$equilibriateNaturalPopulation(params_file, 5.61, 2e6)
```

### Plotting utility

```{r}
source("../R/plot_scan.R")
```

### Simulate

We can now simulate populations with different harvesting rates. Here, we simulate 20 populations with harvesting rates specified by the `hvec` vector. The `simulate_multi` function returns an array if dimensions `{nsteps, length(hvec), 4}`. The last dimension is four because four utilities are returned: {ssb, yield, employment, profit}.

```{r fig.height=9}
nsteps = 200
hvec = seq(0, 0.99, length.out = 20)
lminvec = seq(10, 180, length.out = 20)
Tvec = seq(4, 8, length.out = 20)
lf = 45
h_sq = 0.3
```


### Effect of Min size limit (Dankel et al. model)

Generate reference 

```{r}
dat_ref = data.frame(matrix(ncol=7, nrow=0))
colnames(dat_ref) = c("ssb", "tsb", "yield", "employment.sea", "employment.shore", "profit.sea", "profit.shore")
for (il in 1:17){
  res = simulate(h_sq, lminvec[il], F)
  v = colMeans((res$summaries %>% select(SSB, TSB, Y, D.sea, D.shr, P.sea, P.shr))[101:200,])
  dat_ref[nrow(dat_ref)+1,] = v
}
```

We look at the predictions of the Dankel et al model implementation, and compare it with their results. 

```{r fig.height=9}
fish = new(Fish, params_file)

pop = new(Population, fish)
pop$readParams(params_file, F)
# pop$par$F_spf = 0
pop$par$recruitmentAge = 1
pop$par$dmax = 30000
pop$set_superFishSize(5e6)
pop$verbose = F
pop$init(1000, 5.61)
pop$noFishingEquilibriate(5.61)

sim = new(Simulator, fish)

sim$equilibriateNaturalPopulation(params_file, 5.61, 2e6)

nsteps = 200
hvec = seq(0.01, 0.99, length.out = 20)
lminvec = seq(10, 180, length.out = 20)
Tvec = seq(4, 8, length.out = 20)

if(re_simulate){
  res_ibm_full = sim$simulate_multi_2d(pop, c(5.61), lminvec, h_sq, nsteps, 1.93e3, F)
  save(res_ibm_full, file="outputs/res_ibm_scan_lmin_Dankel22.Rdata")
} else {
  load("outputs/res_ibm_scan_lmin_Dankel22.Rdata")
}

arr = res_ibm_full 
names = pop$colnames
dat_ibm = data.frame(matrix(ncol=length(names), nrow=0))
colnames(dat_ibm) = names
for (i in 1:length(lminvec)){
  v_ibm = colMeans(arr[101:200, 1, i, 1,])
  dat_ibm[nrow(dat_ibm)+1,] = v_ibm
}

plot_scan(dat_ibm, xname = expression("Min size limit"), xvec = lminvec, dat_ref=dat_ref)
```


### Effect of Min size limit (Bioenergetic model)

```{r fig.height=9}
params_file = "../params/cod_params.ini"

fish = new(Fish, params_file)

pop = new(Population, fish)
pop$readParams(params_file, F)
# pop$par$F_spf = 0
pop$par$recruitmentAge = 1
pop$par$dmax = 30000
pop$set_superFishSize(5e6)
pop$verbose = F
pop$init(1000, 5.61)
pop$noFishingEquilibriate(5.61)

sim = new(Simulator, fish)

sim$equilibriateNaturalPopulation(params_file, 5.61, 2e6)

nsteps = 200
hvec = seq(0.01, 0.99, length.out = 20)
lminvec = seq(10, 180, length.out = 20)
Tvec = seq(4, 8, length.out = 20)

if(re_simulate){
  res_ibm_full = sim$simulate_multi_2d(pop, c(5.61), lminvec, h_sq, nsteps, 1.93e3, F)
  save(res_ibm_full, file="outputs/res_ibm_scan_lmin.Rdata")
} else {
  load("outputs/res_ibm_scan_lmin.Rdata")
}

arr = res_ibm_full 
names = pop$colnames
dat_ibm = data.frame(matrix(ncol=length(names), nrow=0))
colnames(dat_ibm) = names
for (i in 1:length(lminvec)){
  v_ibm = colMeans(arr[101:200, 1, i, 1,])
  dat_ibm[nrow(dat_ibm)+1,] = v_ibm
}

plot_scan(dat_ibm, xname = expression("Min size limit"), xvec = lminvec, dat_ref=dat_ref)

dat_ibm %>% with(plot(chi~lminvec))
```