---
title: "Effort control: fishing until quota is filled - stock implementation"
author: "Jaideep Joshi"
date: "28 October 2024"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(fisheRy)
library(tidyverse)
source(here::here("tests/ref/parameters.cod.R"))
source(here::here("tests/ref/simulator.7.R"))
```


```{r}
params_file = here::here("params/cod_params.ini")
params_file_fleet1 = here::here("params/fleet_2_params.ini")
params_file_fleet2 = here::here("params/fleet_3_params.ini")
```

## Initialize a population

To simulate a fished population, we start with creating a fish and constructing a population from it, as usual.

```{r }
fish = new(Fish, params_file)
fish$setMortalityCurveEmpirical(here::here("data/naturalmort.spline.csv"))

pop = new(Stock, fish)
pop$readParams(params_file, F)
pop$superfish_size = 1e6  # Each superfish contains so many fish

pop$superfish_size

```

## Perform Spinup

For simulating fishing, it is important for the population to start with a realistic state distribution. Therfore, we "spin-up" the population under no-fishing conditions. 

```{r}
v = pop$equilibriate_without_fishing(5.61, 200) |> matrix(nrow=200, byrow=T) |> as.data.frame() |> setNames(c("ssb", "tsb", "maturity", "nr", "dr", "nfish"))
```

Let's see how the population state looks after no-fishing equilibrium.

```{r fig.height=9}
d = pop$get_state()
dist = table(d$age, d$length)

par(mfrow = c(5,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5)
plot(v$nfish~seq(1,200,1), ylab="No. of superfish", xlab="Year", type="l", col=c("cyan3"))
plot(v$maturity~seq(1,200,1), ylab="Maturity", xlab="Year", type="l", col=c("cyan3"))
plot(I(v$nr/1e9)~seq(1,200,1), ylab="Recruitment", xlab="Year", type="l", col=c("cyan3"))
plot(v$dr~seq(1,200,1), ylab="actual/potential", xlab="Year", type="l", col=c("cyan3"))

res = simulate(0, 45, F)
matplot(cbind(v$ssb/1e9, res$summaries$SSB[1:200]/1e9), ylab="SSB (MT)", xlab="Year", col=c("cyan3", "black"), lty=1, type=c("p","l"), pch=1)

```


### Set fishing parameters

Set a harvest proportion that will determine the fishing quota.

```{r}
# selectivity
F1 = 0.6901
F2 = 0.06517
F3 = 68.81
F4 = 0.4151
F5 = 112.8
F6 = 0.4381674
lmin = 45

progress_names = c(
  "age",
  "B",
  "B_sampled",
  "yield",
  "yield_expected",
  "fleet0.chi",
  "fleet0.chi_w",
  "fleet0.B_sampled",
  "fleet0.B_start",
  "fleet0.yield",
  "fleet0.F_fishable",
  "fleet0.effortC",
  "fleet0.effortF",
  "fleet1.chi",
  "fleet1.chi_w",
  "fleet1.B_sampled",
  "fleet1.B_start",
  "fleet1.yield",
  "fleet1.F_fishable",
  "fleet1.effortC",
  "fleet1.effortF"
)

```

```{r}
empirical_F_file = ""
mode = "exp"
slope = 0

fleet1 = new(Fleet)
fleet1$readParams(params_file_fleet1, F) 
if (empirical_F_file == ""){
  fleet1$set_referenceFishingMortalityCurveLogistic(F1,F2,F3,F4,F5,F6,lmin)
} else {
  fleet2$set_referenceFishingMortalityCurveEmpirical(empirical_F_file, lmin)
} 
fleet1$par$control_model = mode
fleet1$par$chi0_scalar_slope = slope

fleet2 = new(Fleet)
fleet2$readParams(params_file_fleet2, F) 
if (empirical_F_file == ""){
  fleet2$set_referenceFishingMortalityCurveLogistic(F1,F2,F3,F4,F5,F6,lmin)
} else {
  fleet2$set_referenceFishingMortalityCurveEmpirical(empirical_F_file, lmin)
} 
fleet2$par$control_model = mode
fleet2$par$chi0_scalar_slope = slope

k_unfished = fleet1$biomassFishable(pop, 0, FALSE)
quota = 0.3 * k_unfished
F_fgf = -log(1-0.3)



fleet1$init_chi(pop, F_fgf*0.3, 5.61)
fleet2$init_chi(pop, F_fgf*0.7, 5.61)

fleet1$debug = F
fleet2$debug = F

out = get_fished_dry_run_wrapper(pop, list(fleet1, fleet2), c(quota*0.3, quota*0.7), 5.61, TRUE, TRUE)
df = as.data.frame(matrix(data = out, ncol=length(progress_names), byrow = T))
colnames(df) = progress_names

p = df %>% 
  dplyr::select(-fleet1.chi_w, -fleet0.chi_w) %>%
  mutate(fish = 1:n()) %>% 
  mutate(yield_err_pc = (yield-yield_expected)/yield_expected*100) %>%
  select(-B) %>% 
  pivot_longer(-fish) %>% 
  ggplot(aes(x=fish, y=value))+
  geom_line()+
  facet_wrap(~name, scales="free_y")

print(p)
```

## Calculate error characteristics with 2 fleets

```{r}
pc_error = function(i, mode, h){
  # quota = h*ssb_nf

  fleet1$par$control_model = mode
  fleet2$par$control_model = mode

  k_unfished = fleet1$biomassFishable(pop, 0, FALSE)
  quota = h * k_unfished
  F_fgf = -log(1-h)

  fleet1$init_chi(pop, F_fgf*0.3, 5.61)
  fleet2$init_chi(pop, F_fgf*0.7, 5.61)

  out = get_fished_dry_run_wrapper(pop, list(fleet1, fleet2), c(quota*0.3, quota*0.7), 5.61, TRUE, FALSE)

  df = as.data.frame(matrix(data = out, ncol=2, byrow = T))
  colnames(df) = c("yield1", "yield2")

  df %>% 
    mutate(yield = yield1 + yield2,
           err = (yield-quota)/quota*100,
           yield1pc = yield1/yield * 100,
           yield2pc = yield2/yield * 100
           )
  
}
```

### Try exponential model

$Y/B_s = 1-e^{-k \chi}$

```{r}
pc_error(1, mode="exp", h=0.2)
```

```{r}
niterations = 100

res_exp = tibble(i = 1:niterations) %>%
  mutate(res = map(i, ~pc_error(.x, mode="exp", h=0.3))) %>%
  unnest(res)

print(
res_exp %>%
  ggplot(aes(x=err)) +
  geom_histogram(bins=20) +
  labs(title="Exp model", x="Error (%)", y="Frequency")
)
```

```{r}
res = expand.grid(h = seq(0.1, 0.99, length.out=10), i = 1:niterations) |>
  mutate(res = purrr::pmap(list(i, mode="exp", h), pc_error)) |>
  unnest(res)

res |>
  ggplot(aes(x=factor(h), y=yield1/(yield1+yield2))) +
  geom_violin() +
  geom_hline(yintercept = 0.3, col="red")+
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="red") +
  ylim(0,1)+
  labs(x="Harvest rate (h)", y="yield1/total", title="Error distribution (Exp model)")

res |>
  ggplot(aes(x=factor(h), y=err)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="red") +
  labs(x="Harvest rate (h)", y="Error (%)", title="Error distribution (Exp model)")

res |> 
  ggplot(aes(x=yield2, y=yield1)) +
  geom_point()+
  geom_abline(slope = 3/7, intercept = 0, col="red")
```
