---
title: "Effort control: fishing until quota is filled"
author: "Jaideep Joshi"
date: "28 October 2024"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(fisheRy)
library(tidyverse)
source("../tests/ref/parameters.cod.R")
source("../tests/ref/simulator.7.R")
```


```{r}
params_file = "../params/cod_params.ini"
```

## Initialize a population

To simulate a fished population, we start with creating a fish and constructing a population from it, as usual.

```{r }
fish = new(Fish, params_file)

pop = new(Population, fish)
pop$readParams(params_file, F)
pop$par$n = 2e6  # Each superfish contains so many fish

```

## Perform Spinup

For simulating fishing, it is important for the population to start with a realistic state distribution. Therfore, we "spin-up" the population under no-fishing conditions. 

```{r}
state_nf = pop$noFishingEquilibriate(5.61)
ssb_nf = state_nf[1]
```

Let's see how the population state looks after no-fishing equilibrium.

```{r}
d = pop$get_state()
dist = table(d$age, d$length)

image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length")

```


### Set fishing parameters

Set a harvest proportion that will determine the fishing quota.

```{r}
h = 0.99
pop$set_harvestProp(h)
quota = h*ssb_nf

progress_names = c("fish_age",
                   "B",
                   "B_sampled",
                   "yield",
                   "yield_expected",
                   "chi",
                   "chi_window",
                   "b_sampled_window",
                   "b_start_window",
                   "yield_window")
```

## Mock fish the population to calculate fishing mortality rate (linear model)

Using the linear control model

$Y/B_s = k \chi$


```{r}
fleet = new(Fleet)
fleet$control_model = "linear"
fleet$chi = 100
fleet$init_chi(pop, pop$par$Fc, pop$par$rho, 5.61)

out = fleet$harvest_dry_run(pop, quota, 5.61)

df = as.data.frame(matrix(data = out, ncol=length(progress_names), byrow = T))
colnames(df) = progress_names

df %>% 
  mutate(fish = 1:n()) %>% 
  mutate(yield_err_pc = (yield-yield_expected)/yield_expected*100) %>%
  select(-B) %>% 
  pivot_longer(-fish) %>% 
  ggplot(aes(x=fish, y=value))+
  geom_line()+
  facet_wrap(~name, scales="free_y")

table(df$chi)


fleet$effort_constantC(2.83e-6, 0.75, ssb_nf)*pop$par$dsea
fleet$effort_constantF(2.83e-6, 0.75, ssb_nf)*pop$par$dsea

```


Try one more time

```{r}
fleet = new(Fleet)
fleet$control_model = "linear"
fleet$chi = 100
fleet$init_chi(pop, pop$par$Fc, pop$par$rho, 5.61)

out = fleet$harvest_dry_run(pop, quota, 5.61)

df = as.data.frame(matrix(data = out, ncol=length(progress_names), byrow = T))
colnames(df) = progress_names

df %>% 
  mutate(fish = 1:n()) %>% 
  mutate(yield_err_pc = (yield-yield_expected)/yield_expected*100) %>%
  select(-B) %>% 
  pivot_longer(-fish) %>% 
  ggplot(aes(x=fish, y=value))+
  geom_line()+
  facet_wrap(~name, scales="free_y")

table(df$chi)


fleet$effort_constantC(2.83e-6, 0.75, ssb_nf)*pop$par$dsea
fleet$effort_constantF(2.83e-6, 0.75, ssb_nf)*pop$par$dsea

```


## Mock fish the population to calculate fishing mortality rate (exp model)

Using the exponential control model

```{r}
fleet = new(Fleet)
fleet$chi = 100
fleet$control_model = "exp"
fleet$init_chi(pop, pop$par$Fc, pop$par$rho, 5.61)

out = fleet$harvest_dry_run(pop, quota, 5.61)

df = as.data.frame(matrix(data = out, ncol=length(progress_names), byrow = T))
colnames(df) = progress_names

df %>% 
  mutate(fish = 1:n()) %>% 
  mutate(yield_err_pc = (yield-yield_expected)/yield_expected*100) %>%
  select(-B) %>% 
  pivot_longer(-fish) %>% 
  ggplot(aes(x=fish, y=value))+
  geom_line()+
  facet_wrap(~name, scales="free_y")

```

## Calculate error characteristics 

```{r}
pc_error = function(i, mode, h){
  pop$set_harvestProp(h)
  quota = h*ssb_nf

  fleet = new(Fleet)
  fleet$control_model = mode
  fleet$chi = 100
  fleet$init_chi(pop, pop$par$Fc, pop$par$rho, 5.61)
  
  out = fleet$harvest_dry_run(pop, quota, 5.61)
  
  df = as.data.frame(matrix(data = out, ncol=length(progress_names), byrow = T))
  colnames(df) = progress_names
  
  data.frame(
    err = df %>% 
        mutate(yield_err_pc = (yield-yield_expected)/yield_expected*100) %>% 
        pull(yield_err_pc) %>% 
        tail(1),
    effort_c = fleet$effort_constantC(2.83e-6, 0.75, ssb_nf)*pop$par$dsea,
    effort_f = fleet$effort_constantF(2.83e-6, 0.75, ssb_nf)*pop$par$dsea
  )
}
```

### Try exponential model

$Y/B_s = 1-e^{-k \chi}$

```{r}
res_exp = tibble(i = 1:500) %>%
  mutate(res = map(i, ~pc_error(.x, mode="exp", h=h))) %>%
  unnest(res)

res_exp %>%
  ggplot(aes(x=err)) +
  geom_histogram(bins=20) +
  labs(title="Exp model", x="Error (%)", y="Frequency")
```

### Try linear model

$Y/B_s = k \chi$


```{r}
res_linear = tibble(i = 1:500) %>%
  mutate(res = map(i, ~pc_error(.x, mode="linear", h=h))) %>%
  unnest(res)

res_linear %>%
  ggplot(aes(x=err)) +
  geom_histogram(bins=20) +
  labs(title="linear model", x="Error (%)", y="Frequency")
```

```{r}
res = expand.grid(h = seq(0.1, 0.99, length.out=10), i = 1:500) |>
  mutate(res = purrr::pmap(list(i, mode="linear", h), pc_error)) |>
  unnest(res)

res |>
  ggplot(aes(x=factor(h), y=err)) +
  geom_violin() +
  labs(x="Harvest rate (h)", y="Error (%)", title="Error distribution by harvest rate")

res |> 
  ggplot(aes(x=effort_c, y=effort_f)) +
  geom_point()+
  geom_abline(slope = 1, intercept = 0, col="red")
```


