---
title: "Fished Population and Stock Analysis"
author: "Jaideep"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(fisheRy)
```


# Fishery System Setup

Set up the fishery system using parameter files for fleet and fish, and inspect initial parameters.

```{r setup-fishery}
params_file_fleet = here::here("params/fleet_1_params.ini")
params_file_fish  = here::here("params/cod_params.ini")

fleet = new(Fleet)
fleet$readParams(params_file_fleet, T)
fleet$par$print();

fish = new(Fish, params_file_fish)

fishery = new(Fishery, params_file_fish, fish);
fishery$par$print();
```

# Fishery Simulation

Configure the fishery, set harvest proportion, add fleet, and run equilibrium and initialization steps.

```{r configure-fishery}
fishery$set_harvestProp(0.5);
fishery$addFleet(params_file_fleet, T);

fishery$equilibriateNaturalPopulation(5.61, 2e6);

fishery$init(1000, 0, 5.61);

quota = fishery$calc_quota(5.61);
cat("Quota: ", quota, '\n')
```

# Running and Visualizing Simulation

Simulate the fishery over time and visualize key outputs.

```{r run-simulation}
dat <- fishery$simulate(45, 0.55, 300, 0, 5.61, T, here::here("fishery_output/age_dist_pred.csv"))
dat |> head()
```

## Plot all output variables

```{r}
dat |> 
  mutate(t=1:n()) |>
  pivot_longer(-t) |> 
  ggplot(aes(x=t, y=value)) +
  geom_line(col="seagreen") +
  facet_wrap(~name, scales="free_y")
  
```

## Quota (FGF) vs Yield

This plot compares the feeding ground fishery quota to the yield, excluding the first 10 iterations.

```{r plot-quota-fgf-yield}
dat |> 
  slice(-(1:10)) |>
  ggplot(aes(x=quota_fgf, y=yield)) +
  geom_point() +
  geom_abline(slope=1, intercept=0, color="red")
```

## TSB vs Quota (FGF)

This plot shows the relationship between total stock biomass and feeding ground fishery quota. Slope should be equal to harvest proportion.

```{r plot-tsb-quota-fgf}
dat |> 
  slice(-(1:10)) |>
  ggplot(aes(x=tsb, y=quota_fgf)) +
  geom_point() +
  geom_smooth(method="lm", se=F, formula = y ~ x-1) +
  geom_abline(slope=0.5, intercept=0, color="red")
```

## SSB Over Simulation

This plot tracks spawning stock biomass over the course of the simulation.

```{r plot-ssb-simulation}
dat |> 
  ggplot(aes(x=1:nrow(dat), y=ssb/1e9)) +
  geom_line()
```

