---
title: "Fished Population and Stock Analysis"
author: "Jaideep"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(fisheRy)

setwd(here::here("vignettes"))
```


# Fishery System Setup

To simulate a fishery, i.e., a fished population, we need a `Stock` to harvest, a `Fleet` that harvests it, and a `Fishery` that manages this interaction. A Fishery also contains a hypothetical unfished population, separate from the actual population that we will fish.

Fish and population parameters are the same as in the case of an unfished population.
Fleet parameters must be defined in a separate parameters file. Each fleet needs its own parameters file.

Let us now set up the fishery system using parameter files for fleet and fish, and inspect initial parameters.

```{r setup-fishery}
# define parameter files
params_file_fleet = here::here("params/fleet_1_params.ini")
params_file_spf = here::here("params/fleet_1s_params.ini")
params_file_fish  = here::here("params/cod_params.ini")

# Create a prototype fish to construct the population
fish = new(Fish, params_file_fish)
# fish$natural_mort_scalar = 1
# fish$setMortalityCurveEmpirical(here::here("data/naturalmort.spline.csv"))
fish$par$print()

# Craeate the fishery. Note that we do not need to explicitly create a population - it is contained within the fishery
fishery = new(Fishery, params_file_fish, fish);
fishery$par$print()
```

# Fishery Simulation

Configure the fishery:

- set contol parameters, such as harvest proportion and min size limit
- add the fleet to the fishery
- A fishery simulation requires an estimate of the carrying capacity of the ocean, which is computed by simulating a hypothetical unfished population to equilibrium. This unfished population is also implicitly included in the Fishery class.
- Initialize the fishery - this creates fish in the real fish population.

```{r configure-fishery}
h = 0.22
fishery$set_harvestProp(h);
fishery$addFleet(params_file_fleet, T);
fishery$addSpawnerFleet(params_file_spf, T);

# Verify added fleet's fishing mortality curve
tibble(len=seq(10,200,length.out=100)) |> 
  mutate(Fref = purrr::map_dbl(len, ~fishery$get_fref(0, .x))) |> 
  ggplot(aes(x=len, y=Fref))+
  geom_line()

```

```{r}
# Simulate the Fishery's hypothetical unfished population to equilibrium
v = fishery$equilibriateNaturalPopulation(5.61, 2e6, 200);
fishery$set_superFishSize(2e6)

# Initialize the fishery and equilibriate without fishing
fishery$init(1000, 0, 5.61);
v2 = fishery$equilibriateWithoutFishing(5.61, 200)

# Equilibriate under status quo mortality
v3 <- fishery$simulate(45, 0.22, 200, 1.93e3, rep(5.61, 200), rep(1, 200), F, here::here("fishery_output/age_dist_pred.csv"))

# Set desired total recruitment noise (ideally from params file)
rec_noise_sd = 0.4858775

# Calculate CV of environmental noise, avoid negatives
cv_rec_noise_demographic = sd(v3$recruits)/mean(v3$recruits)
rec_noise_sd = max(sqrt(rec_noise_sd^2 - cv_rec_noise_demographic^2), 0)

# Generate recruitment noise multipliers
rec_noise_t = exp(rnorm(n=300, mean=-rec_noise_sd*rec_noise_sd/2, sd=rec_noise_sd))

# Just a demo calc of the initial quota corresponding to the set defined harvest proportion. The Fishery will recalculate this every year.
quota = fishery$calc_quota(5.61);
cat("Quota: ", quota, '\n')
```

# Running and Visualizing Simulation

Simulate the fishery over time and visualize key outputs.

```{r run-simulation}
# Visualize recruitment noise multipliers
hist(rec_noise_t, 20)

# Run actual run under desired control params
dat <- fishery$simulate(45, 0.5, 300, 1.93e3, rep(5.61, 300), rec_noise_t, F, here::here("fishery_output/age_dist_pred.csv"))
dat |> head()
```
\## Plot all output variables

```{r plot-all}
p1 <- dplyr::bind_rows(v3, dat) |>
  mutate(t=1:n()) |>
  pivot_longer(-t) |>
  ggplot(aes(x=t, y=value)) +
  geom_line(col="seagreen") +
  facet_wrap(~name, scales="free_y")

p2 <- dplyr::bind_rows( 
  dat |> select(
    sim = ssb_spawning, ref = ssb_spawning_ref, 
  ) |> 
  mutate(group = "before spawning"),

  dat |> select(
    sim = ssb_after_spawning, ref = ssb_after_spawning_ref, 
  ) |> 
  mutate(group = "after_spawning"),

  dat |> select(
    sim = ssbn, ref = ssbn_ref
  ) |> 
  mutate(group = "after_spawning_&_spf"),
) |> 
  as_tibble() |> 
  ggplot(aes(x=ref, y=sim)) + 
  geom_point() + 
  geom_abline(slope=1,intercept=1,col="red")+
  facet_wrap(~group)

print(p1)
print(p2)
```


## Quota (FGF) vs Yield

This plot compares the feeding ground fishery quota to the yield, excluding the first 10 iterations.

```{r plot-quota-fgf-yield}
dat |> 
  slice(-(1:10)) |>
  ggplot(aes(x=quota, y=yield)) +
  geom_point() +
  geom_abline(slope=1, intercept=0, color="red")

dat |> 
  slice(-(1:10)) |>
  ggplot(aes(x=quota_fgf, y=yield_fgf)) +
  geom_point() +
  geom_abline(slope=1, intercept=0, color="red")

dat |> 
  slice(-(1:10)) |>
  ggplot(aes(x=quota_spf, y=yield_spf)) +
  geom_point() +
  geom_abline(slope=1, intercept=0, color="red")


```

## TSB vs Quota (FGF)

This plot shows the relationship between total stock biomass and feeding ground fishery quota. Slope should be equal to harvest proportion.

```{r plot-tsb-quota-fgf}
dat |> 
  slice(-(1:10)) |>
  ggplot(aes(x=tsb, y=quota)) +
  geom_point() +
  geom_smooth(method="lm", se=F, formula = y ~ x-1) +
  geom_abline(slope=h, intercept=0, color="red")
```

## SSB Over Simulation

This plot tracks spawning stock biomass over the course of the simulation.

```{r plot-ssb-simulation}
dat |> 
  mutate(t = 1:n()) |> 
  # filter(t > 200) |>
  ggplot(aes(x=t, y=ssb/1e9)) +
  geom_line()

dat |> 
  mutate(t = 1:n()) |> 
  filter(t > 200) |>
  ggplot() +
  geom_line(aes(x=t, y=quota_fgf/1e9)) +
  geom_line(aes(x=t, y=yield_fgf/1e9), col="red") 
```

