---
title: "Simulating a fish population with fishing (using Fishery Class)"
author: "Jaideep Joshi"
date: "24 March 2025"
output: html_document
editor_options:
  chunk_output_type: inline
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(fisheRy)
source(here::here("tests/ref/parameters.cod.R"))
source(here::here("tests/ref/simulator.7.R"))
```


```{r}
params_file = here::here("params/cod_params.ini")
```

## Initialize a Fishery

To simulate a fished population, we start with creating a fish and constructing a population from it, as usual.

```{r }
fish = new(Fish, params_file)

fishery = new(Fishery, params_file, fish)
fishery$equilibriateNaturalPopulation(5.6, 2e6)

```

## Perform Spinup

For simulating fishing, it is important for the population to start with a realistic state distribution. Therfore, we "spin-up" the population under no-fishing conditions. 

```{r}
fishery$pop$print_summary()
fishery$pop$init(1000, 5.61)
fishery$pop$print_summary()
fishery$init(1000, 5.61)

state_nf = fishery$pop$noFishingEquilibriate(5.61)
ssb_nf = state_nf[1]
```

Let's see how the population state looks after no-fishing equilibrium.

```{r}
d = fishery$pop$get_state()
dist = table(d$age, d$length)

image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length")

```


### Set fishing parameters

Once we have created a population, we can set the fishing parameters:
  
+ Harvest proportion - the proportion of fish that are allowed to be harvested every year
+ Minimum size limit - the size below which fish cannot be harvested

The most important step is to set the `simulate_bio_only` flag, because simulating the socioeconomic dynamics requires additional steps, which we will show in another vignette.

```{r}
h = 0.3
pop$set_harvestProp(h)   # Harvest 
pop$set_minSizeLimit(45) # Only fish above 45 cm length can be harvested
```

```{r}
pop$par$simulate_bio_only = T # VERY IMPORTANT
```

## Simulate the population

Now we are ready to simulate the population with fishing by continuously `update`ing it. Each update performs 1 year of simulation, implementing the processes of maturation, growth, mortality, and reproduction for each fish in the population.

Let's confirm that the parameters we have set are registered:
```{r}
pop$par$print()
```

This time, we also provide a dynamic temperature timeseries to the population, wherein the first 100 years have a fixed temperature of 5.61 degC, and in the next 100 years, temperature increases linearly to 8 degC.

```{r fig.height=9}
pop$verbose = T
nsteps = 200             # Let's simulate for 200 years
nfish = numeric(nsteps)  # Let's keep track of the number of superfish
nfish[1]=pop$nfish()            
cnames = pop$colnames
dat = data.frame(matrix(ncol=length(cnames), nrow=0))
colnames(dat) = cnames

temp = rep(5.61, nsteps)
temp[(nsteps/2+1):nsteps] = seq(5.61,8, length.out=nsteps/2)

for (i in 1:nsteps){
  v = pop$update(temp[i])       # Update all fish over 1 year 
  dat[i,] = v    # some book-keeping
  nfish[i] = pop$nfish()
}

d = pop$get_state()
dist = table(d$age, d$length)

par(mfrow = c(4,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5)
plot(temp~seq(1,nsteps))
plot(nfish~seq(1,nsteps,1), ylab="No. of superfish", xlab="Year")
image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length")

res = simulate(h, 45, F)
matplot(cbind(dat$ssb/1e9, res$summaries$SSB[1:nsteps]/1e9), ylab="SSB (MT)", xlab="Year", col=c("cyan", "black"), lty=1, type=c("o","l"), pch=1)
abline(h=ssb_nf/1e9, col="grey") 

```
