---
title: "Safe operating spaces"
author: "Jaideep Joshi"
date: "28 March 2022"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(fisheRy)
library(tidyverse)
source("../tests/ref/parameters.cod.R")
source("../tests/ref/simulator.7.R")
```

```{r}
re_simulate = T

# if one of the files doesnt exist, re-run the simulations
re_simulate = re_simulate | !file.exists("outputs/res_ibm_jss_hxL.Rdata")
```

```{r}
params_file = "../params/cod_params.ini"
```


### Simulating the fishery with socioeconomic outputs

The socio-economic part of the fishery deals with the calculation of realized fishing effort for a desired harvest rate, the resulting employment at sea and on shore, and the profit from operating the fishery.

The effort calculations need require the equilibrium fishable biomass of an equivalent unfished population. To simplify such these simulations, fisheRy provides a `Simulator` which can not only precompute the properties of such a  hypothetical unfished population, but also allows simulating multiple populations under different management regimes. 

We start with creating a protoype fish as usual,

```{r}
# Create a prototype fish as usual
fish = new(Fish, params_file)
```

Then we create a simulator and run the equivalent unfished population to equilibrium.

```{r}
sim = new(Simulator, fish)
sim$equilibriateNaturalPopulation(params_file, 5.61, 2e6)
```

```{r}
plot_timeseries = function(res_ibm, pop, h, nsteps=200){
  
  d = pop$get_state()
  table(d$age)
  par(mfrow = c(4,3), mar=c(4,4,1,1))
  
  plot(y=res_ibm$ssb/1e9, x=seq(1,nsteps,1), ylab="Spawning stock biomass (Mt)", xlab="Year", col="cyan3", type="l")
  plot(y=res_ibm$yield/1e9, x=seq(1,nsteps,1), ylab="Yield (Mt)", xlab="Year", col="cyan3", type="l")
  plot(y=res_ibm$employment.sea, x=seq(1,nsteps,1), ylab="Employment at sea (person-years)", xlab="Year", col="cyan3", type="l")
  plot(y=res_ibm$employment.shore, x=seq(1,nsteps,1), ylab="Employment on shore (person-years)", xlab="Year", col="cyan3", type="l")
  plot(y=res_ibm$profit.sea/1e9, x=seq(1,nsteps,1), ylab="Profit at sea (Billions NOK)", xlab="Year", col="cyan3", type="l")
  plot(y=res_ibm$profit.shore/1e9, x=seq(1,nsteps,1), ylab="Profit at shore (Billions NOK)", xlab="Year", col="cyan3", type="l")
  d = pop$get_state()
  d1 = d %>% group_by(age) %>% summarize(mat = length(which(isMature))/length(isMature))
  plot(mat, type="l")
  points(d1$mat~I(d1$age-1), type="o", col="cyan3") # Decrement age to get the right maturation prob (see note above)
    plot(res_ibm$r0~seq(1,nsteps,1), xlab="Year", ylab= "r0", col="cyan3", type="p", ylim=c(0,30))
    abline(h=21.77, col="black")
    matplot(y=cbind(res_ibm$survival_mean, res_ibm$maturity, res_ibm$Nrel), x=seq(1,nsteps,1), xlab="Year", ylab= "p_survival / maturity / Nrel", col=c("cyan3", "magenta3", "yellow3"), type="l", lty=1)
  matplot(y=cbind(res_ibm$factor_dg, res_ibm$factor_dr), x=seq(1,nsteps,1), xlab="Year", ylab= "dl_ratio, r_ratio", col=c("cyan3", "magenta3"), type="l", lty=1)
  matplot(y=cbind(res_ibm$length90, res_ibm$max_length), x=seq(1,nsteps,1), xlab="Year", ylab= "Length (90 %ile)", col=c("cyan3", "cyan4"), type="l", lty=1)
  dist = table(d$age, d$length)
  image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length")
}
```


Then, we create the fished population and simulate it for 500 years with a status quo management regime with 
$h = 0.41$, $L50 = 45$, $T = 5.6$.

We can also plot biological and socioeconomic parameters as a function of time.

```{r}
pop_sq = new(Population, fish)
pop_sq$readParams(params_file, F)
pop_sq$set_superFishSize(2e6)
pop_sq$init(1000, 5.61)
pop_sq$noFishingEquilibriate(5.61)

h_sq = 0.05
lf_sq = 45
t_sq = 5.6
nsteps_sq = 200
res_ibm_sq = sim$simulate(pop_sq, lf_sq, h_sq, nsteps_sq, 1.93e3, t_sq, F, "")

plot_timeseries(res_ibm_sq, pop_sq, h_sq, nsteps_sq)
```

We can continue to simulate this population under different management conditions, e.g., here with a higher harvest proportion of 0.2.

```{r}
nsteps = 50
h = 0.2
lf = 45

pop1 = pop_sq ## NOTE:: This creates a shallow copy, so pop_sq will also be modified.
res_ibm = sim$simulate(pop1, lf, h, nsteps, 1.93e3, 5.61, F, "")

plot_timeseries(res_ibm, pop1, h, nsteps)

utils = cbind(res_ibm$ssb/max(res_ibm$ssb), res_ibm$yield/max(res_ibm$yield), res_ibm$employment/max(res_ibm$employment), res_ibm$profit/max(res_ibm$profit))

par(mfrow=c(1,1), mar=c(4,4,1,1))
matplot(y=utils, x=1:50, xlab="Years", ylab="Utilities", col=c("darkgreen", "darkgoldenrod1", "dodgerblue3", "coral1"), lty=1, lwd=2, type="l")
```

Thus, we can plot the two populations together:

```{r}
plot(y=c(res_ibm_sq$ssb, res_ibm$ssb)/1e9, x=seq(1,nsteps_sq+nsteps), type="l", col="cyan3", xlab="Year", ylab="SSB (MT)", lwd=2)
abline(v=nsteps_sq)
```

### Multiple populations

Computing JSS requires that we analyze the population's response to different management regimes. Here, we simulate 25 populations for 50 years under different harvesting regimes (specified by the `hvec` vector). Each population starts from an equilibrium under the same status quo management regime with $h=0.41$, $L_{min}=45$, and $T=5.61$.

The `simulate_multi_2d` function returns an array of dimensions `{nsteps, length(hvec), length(lmin_vec), length(Tvec), 4}`. The last dimension is four because four utilities are returned: {ssb, yield, employment, profit}.

```{r}
source("../R/plot_scan.R")
```

```{r fig.height=9}

# Recreate status quo population because the earlier one was modified 
pop_sq = new(Population, fish)
pop_sq$readParams(params_file, F)
pop_sq$set_superFishSize(2e6)
pop_sq$init(1000, 5.61)
pop_sq$noFishingEquilibriate(5.61)

h_sq = 0.41
lf_sq = 45
t_sq = 5.6
nsteps_sq = 100
res_ibm_sq = sim$simulate(pop_sq, lf_sq, h_sq, nsteps_sq, 1.93e3, t_sq, F, "")


nsteps_jss=50
hvec = seq(0., 0.99, length.out=25)
lfvec = seq(10, 150, length.out=25)
tvec = c(5.61)

pop = pop_sq # Remember, this just creates a reference to pop_sq
if (re_simulate){
  pop$set_superFishSize(2e6)
  res_ibm_full = sim$simulate_multi_2d(pop, tvec, lfvec, hvec, nsteps_jss, 1.93e3, F)
  save(res_ibm_full, file = "outputs/res_ibm_jss_hxL.Rdata")
} else{
  load("outputs/res_ibm_jss_hxL.Rdata")
}

arr = res_ibm_full 
names = pop$colnames
dat_ibm = data.frame(matrix(ncol=length(names), nrow=0))
colnames(dat_ibm) = names
for (i in 1:length(hvec)){
  v_ibm = colMeans(arr[, i, 1, 1,])
  dat_ibm[nrow(dat_ibm)+1,] = v_ibm
}

# plot_scan(dat_ibm, xname = "Harvest rate", xvec = hvec)
```

```{r}
# Wrapper for image to use gray for NA
my.image <- function(x,y,z,  zlim, col, na.color='gray95', outside.color='white', ...)
{
  newz.na <- zlim[2]+(zlim[2]-zlim[1])/length(col) # new z for NA
  newz.outside <- zlim[2]+2*(zlim[2]-zlim[1])/length(col) # new z for values outside zlim

  z[which(is.na(z>zlim[2]))] <- newz.na # we affect newz.outside
#  z[which(z<zlim[1] | z>zlim[2])] <- newz.outside # same for newz.na


  zlim[2] <- zlim[2]+2*(zlim[2]-zlim[1])/length(col) # we finally extend the z limits to include the two new values 

  col <- c(col, na.color, outside.color) # we construct the new color range by including: na.color and outside.color

  image(x,y,z,  zlim=zlim, col=col, useRaster = T, ...) # we finally call image(...)
}


# Function to plot color bar
color.bar <- function(lut, min, max=-min, nticks=11, ticks=seq(min, max, len=nticks), title='') {
  scale = (length(lut)-1)/(max-min)
  
  plot(c(0,10), c(min,max), type='n', bty='n', xaxt='n', xlab='', yaxt='n', ylab='', main=title)
  axis(2, ticks, las=1)
  for (i in 1:(length(lut)-1)) {
    y = (i-1)/scale + min
    rect(0,y,10,y+1/scale, col=lut[i], border=NA)
  }
}


```

### Utilties

```{r fig.wigth=9, fig.height=7.5}
arr = res_ibm_full[,,,,1:4]

d = sim$max_avg_utils_2d(c(4,length(tvec), length(lfvec),length(hvec),nsteps_jss), arr)

utils = array(data=d, dim=c(length(hvec),length(lfvec),length(tvec), 4))

for (i in 1:4){
  utils[,,,i][which(utils[,,,4]<0)]=NA
}

par(mfrow=c(2,2), mar=c(5,5,4,1), oma=c(1,1,1,1), cex.lab=1.2)
for (i in 1:4){
  my.image(x=hvec, y=lfvec, z=utils[,,1,i], col=scales::viridis_pal()(100), zlim=c(0,1), main=c("Spawning stock biomass","Yield","Employment","Profit")[i], 
        xlab="Harvest proportion", ylab=expression("Minimum size limit (L"[50]*")"))
}



```

### Stakeholder satisfaction

```{r fig.width=9, fig.height=6.5}
arr = res_ibm_full[,,,,1:4]

ss = sim$stakeholder_satisfaction_2d(c(4,length(tvec), length(lfvec),length(hvec),nsteps_jss), arr)
scs = array(data=ss, dim=c(length(hvec),length(lfvec),length(tvec), 5))

for (i in 1:5){
  scs[,,,i][which(is.na(utils[,,,4]))]=NA
}

par(mfrow=c(2,3), mar=c(5,5,4,1), oma=c(1,1,1,1), cex.lab=1.2)
for (i in 1:5){
  my.image(x=hvec, y=lfvec, z=scs[,,1,i], col=scales::viridis_pal()(100), zlim=c(0,1), main=c("Industrial fishers", "Artisanal fishers", "Employment-prioritizing\npolicymakers", "Profit-prioritizing\npolicymakers", "Conservationists")[i], 
        xlab="Harvest proportion", ylab=expression("Minimum size limit (L"[50]*")"))
}

```

### JSS

```{r fig.wigth=9, fig.height=7.5}

JSS = array(dim=c(length(hvec),length(lfvec),length(tvec),3))

JSS[,,,1] = apply(scs, c(1,2,3), min)
JSS[,,,2] = apply(scs, c(1,2,3), mean)
JSS[,,,3] = 1/apply(1/scs, c(1,2,3), mean)

for (i in 1:3){
  scs[,,,i][which(is.na(utils[,,,4]))]=NA
}

par(mfrow=c(2,2), mar=c(5,5,4,1), oma=c(1,1,1,1), cex.lab=1.2)
cols = scales::viridis_pal()(100)
for (i in 1:3){
  my.image(x=hvec, y=lfvec, z=JSS[,,1,i], col=cols, zlim=c(0,1), main=c("JSS (arithmatic mean)","JSS (harmonic mean)","JSS (minimum)")[i], 
        xlab="Harvest proportion", ylab=expression("Minimum size limit (L"[50]*")"))
}
color.bar(cols, 0,1, nticks = 5)

```


