---
title: 'Fisheries model theory: biological model'
author: "Jaideep Joshi"
date: "8 April 2026"
output:
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

```{r, include = F, echo = F}
library(fisheRy)
library(tidyverse)
source(here::here("tests/ref/parameters.cod.R"))
source(here::here("tests/ref/simulator.7.R"))
```

## Introduction

A fish goes through four life-history processes every year: 

1. Growth - This is the increment in the body size by growth. Growth rates are different for mature and immature individuals, since mature individuals allocate resources for reproduction.

2. Decision to mature - Every year, fish will decide whether or not to mature (until maturation). Once matured, reproduction begins.

3. Reproduction - production of eggs

4. Mortality - This is natural and fishing-induced death 


### Simulating growth


```{r}
params_file = here::here("params/cod_params.ini")
fleet_params_file = here::here("params/fleet_1_params.ini")
```

Let us now simulate the growth of a fish starting at age 1 upto age 30 under constant temperature and TSB. The fish will perform maturation, growth, and age increment, every year in that order. We will record the fish length at each timestep so that we can plot the growth trajectory.

```{r}
fish = new(Fish, params_file)
fish$init(0, 5.61)

years = 1:31
length = numeric(31)
weight = numeric(31)
dw = numeric(31)
gsi = numeric(31)
dl_real = numeric(31)
dl_real_s = numeric(31)
for (i in years){
  length[i] = fish$length
  weight[i] = fish$weight
  fish$updateMaturity(5.61)
  fish$grow(0, 5.61)
  dw[i] = fish$delta_weight
  gsi[i] = fish$gsi_effective
  dl_real[i] = fish$dl_real
  dl_real_s[i] = fish$dl_real_stochastic
  fish$set_age(fish$age+1)
}

tibble(years, length, weight, dw, gsi, dl_factor = dl_real_s/dl_real) |>
  pivot_longer(-years) |> 
  ggplot(aes(x=years, y=value))+
  geom_point()+
  geom_line()+
  facet_wrap(~name, scales="free")

```


Now, since the growth trajectory of a fish depends on the age at which it matures, and since maturation is a random process, let us simulate lots of fish and look at the possible growth trajectories. 

```{r fig.width=4, fig.height=4}
growth_trajectories = function(beta1=NULL, beta2=NULL, N=50, temp_sd = 3, tsb_min = 0, tsb_max=1.93e3, growth_noise_sd = NULL){
  names = c("t_birth",	"age",	"isMature",	"isAlive",	"length",	"weight", "mort", "temp", "fec", "ssb")
  dat_full = data.frame(data=matrix(nrow=0, ncol=length(names)))
  colnames(dat_full) = names
  plot(x=1, y=NA, xlim=c(0,31), ylim=c(0,300), xlab="Age", ylab = "Length")
  for (f in 1:N){
    fish = new(Fish, params_file)
    if (!is.null(growth_noise_sd)) fish$par$growth_noise_sd = growth_noise_sd
    if (!is.null(beta1)) fish$par$beta1 = beta1
    if (!is.null(beta2)) fish$par$beta2 = beta2
  
    dat = data.frame(data=matrix(nrow=0, ncol=length(names)))
    colnames(dat) = names
  
    temp0 = rnorm(1, mean=5.61, sd=temp_sd)
    tsb0 = runif(1, min = tsb_min, max = tsb_max)
    fish$init(tsb0, temp0)
    dat[1,] = c(fish$get_state(), fish$naturalMortalityRate(5.61), 5.61, 0, 0.8*tsb0) # Assuming SSB is 80% of TSB, since we are not simulating a population
    
    for (i in 2:30){
      temp = rnorm(1, mean=5.61, sd=temp_sd)
      tsb = runif(1, min = tsb_min, max = tsb_max)
      fish$updateMaturity(temp)
      fish$grow(tsb, temp)
      fish$set_age(fish$age+1)
      dat[i,] = c(fish$get_state(), fish$naturalMortalityRate(temp), temp, fish$produceRecruits(0.8*tsb*1e6, temp), 0.8*tsb) # Assuming SSB is 80% of TSB, since we are not simulating a population
    }
    points(dat$length~dat$age, type="l", col=scales::alpha(rainbow(N)[f], 0.3))
    dat_full = rbind(dat_full, dat)
  }
  dat_full
}
```

### Growth trajectories with fixed temp, variable TSB (+density effect), and no growth stochasticity

```{r}
dat_full1 = growth_trajectories(N=50, temp_sd=0, growth_noise_sd = 0)

dat_full1 |>
  filter(age <= 15) |>
  ggplot(aes(x=length, group=age, color=age))+
  geom_density()
```

### Growth trajectories with fixed temp, variable TSB (+density effect), and + growth stochasticity

```{r}
dat_full2 = growth_trajectories(N=50, temp_sd=0, growth_noise_sd = 0.5)

dat_full2 |>
  filter(age <= 15) |>
  ggplot(aes(x=length, group=age, color=age))+
  geom_density()
```


```{r}
dat_full3 = growth_trajectories(N=50, temp_sd=0, growth_noise_sd = 0.2347)

dat_full3 |>
  filter(age <= 15) |>
  ggplot(aes(x=length, group=age, color=age))+
  geom_density()
```


```{r}
dat_full = growth_trajectories(N=50, growth_noise_sd = 0.2347)

dat_full |>
  filter(age <= 15) |>
  ggplot(aes(x=length, group=age, color=age))+
  geom_density()
```


## Decision to mature

Decision to mature is defined by a probabilistic maturation reaction norm (PMRN). Let us visualize the PMRN.

```{r}
lvec = seq(0,200, length.out = 100)
avec = seq(1,31, length.out = 31)

fish = new(Fish, params_file)
fish$init(0, 5.61)

calc_maturation_prob1 = function(a,l, temp=5.61){
  fish$set_age(a)
  fish$set_length(l)
  p = fish$maturationProb(temp)
  p
}


list(lvec = seq(0,200, length.out = 100), avec = seq(1,31, length.out = 31)) %>% cross_df() %>% mutate(prob = purrr::map2_dbl(avec, lvec, ~calc_maturation_prob1(.x, .y))) %>% 
ggplot() + 
geom_raster(aes(x=avec, y=lvec, fill=prob)) + scale_fill_viridis_c() + labs(x="Age", y="Length") + theme_classic(base_size=20)


list(lvec = seq(0,200, length.out = 100), temp = c(4.61,5.61,6.61)) %>% cross_df() %>% mutate(prob = purrr::map2_dbl(lvec, temp, ~calc_maturation_prob1(a=10, l=.x, temp = .y))) %>% 
ggplot() + 
geom_line(aes(x=lvec, y=prob, group=temp, col=temp)) + scale_colour_viridis_c() + labs(x="Length", y="Maturation probability") + theme_classic(base_size=20)

```

This PMRN influences the growth trajectories by determining the age at maturation, like so,

```{r}
dat_full_00 = growth_trajectories(0,0)
```

and leads to the following maturity ogives.

```{r fig.width=4, fig.height=5}
par(mfrow = c(2,1), mar=c(4,4,1,1))

dat_full %>% group_by(age) %>% summarize(maturity = mean(isMature)) %>% with(plot(maturity~age, ylab="Maturity", xlab="Age", type="l", col="cyan3", lwd=2))

dat_full %>% mutate(length_class = cut(length, breaks=20)) %>% group_by(length_class) %>% summarize(maturity = mean(isMature), length=mean(length)) %>% with(plot(maturity~length, ylab="Maturity", xlab="Length", type="l", col="cyan3", lwd=2))
```


## Reproduction and recruitment

Cumulative reproductive investment translates into production of eggs. 
```{r}
get_recruits = function(ssb, temp){
  fish = new(Fish, params_file)
  # fish$par$growth_noise_sd = 0
  fish$init(0, temp)
  for (i in 1:15){
    fish$updateMaturity(temp*100)
    fish$grow(ssb/1e6, temp)
    fish$set_age(fish$age+1)
    # print(fish$length)
  }
  fish$print()
  fish$produceRecruits(ssb, temp)
}

list(ssb = seq(0, 5e9, length.out=100), temp = c(4.61, 5.61, 6.61)) %>% 
  cross_df() %>% 
  mutate(eggs = purrr::map2_dbl(ssb, temp, ~get_recruits(.x, .y) )) %>% 
  mutate(eggs0 = purrr::map2_dbl(0, temp, ~get_recruits(.x, .y) )) %>% 
  ggplot(aes(y=eggs/eggs0,x=ssb, col=temp, group=temp)) + 
  geom_point(size=2) + 
  labs(x="Spawning stock biomass (Mt)", y = "Recruits / Recruits_0") + 
  theme_classic(base_size = 20) + 
  scale_colour_viridis_c(end=0.9)
```

Here is the number of recruits per fish as a function of spawning stock biomass obtained from the growth trajectories above.

```{r}
dat_full %>% ggplot(aes(x=ssb, y=fec)) + geom_point(aes(col=temp)) + scale_colour_viridis_c() + labs(x="Spawning stock biomass (kT)", y = "Recruits per capita") + theme_classic(base_size=20) + scale_y_log10(limits = c(1e-4, NA)) 
```

```{r}
dat_full %>% mutate(ssb_class = ave(ssb, cut(ssb, breaks = 30))) %>% group_by(age, ssb_class) %>% summarize(recr = mean(fec)) %>% 
  ggplot() + geom_tile(aes(x=age, y=ssb_class, fill=log(1+recr))) + scale_fill_viridis_c() + labs(y="Spawning stock biomass (kT)", x = "Age") + theme_classic(base_size=20)
```

## Mortality


```{r fig.height=5, fig.width=6}
dat_full %>% 
  ggplot(aes(y=mort, x=length, col=temp))+
  geom_point(aes(shape=factor(isMature)), alpha=0.5)+
  geom_jitter()+
  labs(x="Length", y=expression("Inst. natural mortality rate (yr"^"-1"*")"), col="Temp", )+
  theme_classic(base_size = 20)+
  scale_color_viridis_c()+
  scale_y_log10()+
  scale_x_log10()#+
  #facet_grid(cols=vars(isMature), labeller = as_labeller(c(`0`="Immature", `1`="Mature")))
```

<!-- 
Instantaneous fishing mortality rate depends on the harvest proportion $h$ and fishing selectivity $\sigma$, and acts on top of natural mortality. Thus, for a given harvest proportion $h$ and minimum size limit $L_{50}$, fishing mortality rate is given by 

$$
F = -\log(1-h)\cdot \sigma(l_a),
$$

where $\sigma(l_a)$ is the fishing selectivity,

$$
\sigma(l_a) = \frac{F_1}{1+e^{-F_2(l_a-L_{50})}} - \frac{F_6}{1+e^{-F_4(l_a-F_5)}},
$$

where $s$ is the Slope of the fishing selectivity curve.

```{r}
fleet = new(Fleet)
fleet$readParams(fleet_params_file, F)
fleet$par$print()

fleet2 = new(Fleet)
fleet2$readParams(fleet_params_file, F)
# fleet2$set_minSizeLimit(65)

bind_rows(
  tibble(length = seq(0,200,length.out=100)) %>%
    mutate(mort = purrr::map_dbl(length, ~fleet$fishingMortalityRef(.x))  ) |> 
    mutate(lmin = 45),
  tibble(length = seq(0,200,length.out=100)) %>%
    mutate(mort = purrr::map_dbl(length, ~fleet2$fishingMortalityRef(.x)) ) |>
    mutate(lmin = 65)
  ) %>% 
  ggplot(aes(y=mort, x=length, group=lmin, col=factor(lmin)))+
  geom_line(size=2)+
  labs(x="Length", y="Ref fishing mortality rate", col="Lmin")+
  theme_classic(base_size = 20)


dat_full %>% 
  mutate(sel = purrr::map_dbl(length, ~fleet$fishingMortalityRef(.x))  ) |> 
  ggplot(aes(y=mort-log(1-0.4)*sel, x=length, col=temp))+geom_point(alpha=0.5)+geom_jitter()+labs(x="Length", y=expression("Inst. total mortality rate (yr"^"-1"*")"), col="Temp")+theme_classic(base_size = 20)+scale_color_viridis_c()+scale_y_log10()+scale_x_log10()
```

### Survival probability

Survival probability thus given by the following equation:

$$
s(a,l_a) = e^{- (M(a,l_a) + F)},
$$
where $F$ is the fishing mortality rate.


 -->
