---
title: "Simulating a fish population without fishing"
author: "Jaideep Joshi"
date: "13 Jan 2024"
output: html_document
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = F, echo = F}
library(fisheRy)
source(here::here("tests/ref/parameters.cod.R"))
source(here::here("tests/ref/simulator.7.R"))
```

### Create a fish

A population is a collection of fish. Therefore to create a population, we first create a fish. This fish will be used a prototype to construct the population.

Here, we create a fish and set some parameters:

```{r}
params_file = here::here("params/cod_params_dankel22.ini")
```

Let us set parameters that allow us to reproduce the Dankel et al XXXX model.

```{r }
fish = new(Fish, params_file)
fish$par$Bhalf = 3.65e8

fish$par$growth_model_name = "Dankel22"
fish$par$maturation_model_name = "Dankel22"
fish$par$mortality_model_name = "Dankel22"
fish$par$recruitment_model_name = "BevertonHoltDirect"

fish$init(1.93e3, 5.61)
fish$par$print()
fish$get_state()
```

### Create a population

The constructor for the population requires a Fish object as an argument. This fish object will be used as a prototype to construct all fish in the population. You can create a fish population as follows:

```{r}
pop = new(Stock, fish)  
pop$readParams(params_file, F)
```

### Setting population parameters

Population parameters can be set in the same way as fish parameters, using the `par` object in the population

Two parameters are important:

+ `n` - Superfish size. Since the population can have billions of fish, we do not simulate each individual fish. Rather, each "Fish" represents a group of `n` actual fish. This collection is called "Superfish". Reducing this parameter will necessitate simulating more agents in the model, which will reduce noise but also take more computational time. 
+ `Bhalf` - This is a parameter that controls density dependent recruitment. 


```{r}
pop$superfish_size = 2e6  # Each superfish contains so many fish
```

We can also set the management parameters for a population:

+ Harvest proportion - the proportion of fish that are allowed to be harvested every year
+ Minimum size limit - the size below which fish cannot be harvested

```{r}
# pop$set_harvestProp(0)   # Harvest 
# pop$set_minSizeLimit(45) # Only fish above 45 cm length can be harvested
```

### Initialising a population

By default, a population contains no fish. We must initialize the population with specified number of superfish.

```{r}
pop$init(1000, 0, 5.61)  # initialize the population with 1000 agents (superfish)
```

We can also peek at the state of the population. The state of the population is simply a dataframe consisting of the state of each superfish in the population.

```{r}
d = pop$get_state()
head(d)
```

### Simulating an unfished population

We can simulate a population by continuously `update`ing it. Each update performs 1 year of simulation, implementing the processes of maturation, growth, mortality, and reproduction for each fish in the population.

For example, let is simulate a population for 200 years.

```{r fig.height=9}
# pop$verbose = T
# nsteps = 200             # Let's simulate for 200 years
# nfish = numeric(nsteps)  # Let's keep track of the number of superfish
# nfish[1]=1000            # Since we initialized the population with 1000 superfish
# cnames = pop$colnames
# dat = data.frame(matrix(ncol=length(cnames), nrow=0))
# colnames(dat) = cnames
# for (i in 1:nsteps){
#   v = pop$update(5.61)       # Update all fish over 1 year 
#   dat[i,] = v    # some book-keeping
#   nfish[i] = pop$nfish()
# }

v = pop$equilibriate_debug(5.61) |> matrix(nrow=200, byrow=T) |> as.data.frame() |> setNames(c("ssb", "tsb", "maturity", "nr", "dr", "nfish"))

d = pop$get_state()
dist = table(d$age, d$length)

par(mfrow = c(5,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5)
plot(v$nfish~seq(1,200,1), ylab="No. of superfish", xlab="Year", type="l", col=c("cyan3"))
plot(v$maturity~seq(1,200,1), ylab="Maturity", xlab="Year", type="l", col=c("cyan3"))
plot(I(v$nr/1e9)~seq(1,200,1), ylab="Recruitment", xlab="Year", type="l", col=c("cyan3"))
plot(v$dr~seq(1,200,1), ylab="actual/potential", xlab="Year", type="l", col=c("cyan3"))

res = simulate(0, 45, F)
matplot(cbind(v$ssb/1e9, res$summaries$SSB[1:200]/1e9), ylab="SSB (MT)", xlab="Year", col=c("cyan3", "black"), lty=1, type=c("p","l"), pch=1)
```

```{r fig.height=5}

par(mfrow = c(1,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5)
image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(1+dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length")


```

<!-- #### Let us validate the SSB numbers at various stages of the year -->

<!-- ```{r} -->
<!-- validate_ssb = function(dat, F_spf, f_spf_before){ -->
<!--   # par(mfrow=c(2,2), mar=c(4,4,1,1), oma=c(1,1,1,1)) -->
<!--   # plot(dat$ssb_spawning~dat$ssb_spawning_ref) -->
<!--   # abline(0,1, col="red") -->
<!--   # plot(dat$ssb_after_spawning~dat$ssb_after_spawning_ref) -->
<!--   # abline(0,1, col="red") -->
<!--   # plot(dat$ssbn~dat$ssbn_ref) -->
<!--   # abline(0,1, col="red") -->
<!--   # plot(dat$yield_spf~dat$yield_spf_ref) -->
<!--   # abline(0,1, col="red") -->

<!--   h_spf = 1-exp(-F_spf*1) -->
<!--   p_survival_spf_before = 1 - f_spf_before*h_spf -->
<!--   p_survival_spawning = exp(-fish$par$Mspawning) -->
<!--   p_survival_spf_after = (1-h_spf)/(1 - f_spf_before*h_spf) -->
<!--   p_yield_spf = h_spf*(f_spf_before + (1-f_spf_before)*exp(-fish$par$Mspawning)) -->

<!--   to_seabed_spawning = dat$ssb_spawning-dat$ssb_after_spawning -->

<!--   par(mfrow=c(3,2), mar=c(4,4,1,1), oma=c(1,1,1,1)) -->
<!--   plot(dat$ssb~dat$ssb_spawning) -->
<!--   abline(0,1, col="grey") -->
<!--   abline(0,1/p_survival_spf_before, col="red") -->
<!--   plot(dat$ssb_spawning~dat$ssb_after_spawning) -->
<!--   abline(0,1, col="grey") -->
<!--   abline(0,1/p_survival_spawning, col="red") -->
<!--   plot(dat$ssb_after_spawning~dat$ssbn) -->
<!--   abline(0,1, col="grey") -->
<!--   abline(0,1/p_survival_spf_after, col="red") -->
<!--   plot(I(dat$ssb-dat$ssbn)~I(dat$yield_spf + to_seabed_spawning)) -->
<!--   abline(0,1, col="red") -->

<!--   plot(I(dat$tsb_before_mort - dat$tsb_after_mort)~I(dat$yield+dat$to_sea_bed)) -->
<!--   abline(0,1, col="red") -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- validate_ssb(dat, 0, pop$par$f_spf_before) -->
<!-- ``` -->

<!-- ### Simulate the same model with SPF -->

<!-- ```{r} -->
<!-- fish = new(Fish, params_file) -->
<!-- fish$par$s0 = 0.059 -->

<!-- pop = new(Population, fish)   -->
<!-- pop$readParams(params_file, F) -->
<!-- pop$par$n = 1e6  # Each superfish contains so many fish -->
<!-- pop$par$F_spf = 1.0 -->


<!-- pop$init(1000, 1.93e3, 5.61)  # initialize the population with 1000 agents (superfish) -->

<!-- pop$verbose = T -->
<!-- nsteps = 200             # Let's simulate for 200 years -->
<!-- nfish = numeric(nsteps)  # Let's keep track of the number of superfish -->
<!-- nfish[1]=1000            # Since we initialized the population with 1000 superfish -->
<!-- cnames = pop$colnames -->
<!-- dat = data.frame(matrix(ncol=length(cnames), nrow=0)) -->
<!-- colnames(dat) = cnames -->
<!-- for (i in 1:nsteps){ -->
<!--   v = pop$update(5.61)       # Update all fish over 1 year  -->
<!--   dat[i,] = v    # some book-keeping -->
<!--   nfish[i] = pop$nfish() -->
<!-- } -->

<!-- d = pop$get_state() -->
<!-- dist = table(d$age, d$length) -->

<!-- par(mfrow = c(3,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5) -->
<!-- plot(nfish~seq(1,nsteps,1), ylab="No. of superfish", xlab="Year") -->
<!-- image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length") -->

<!-- res = simulate(0, 45, F) -->
<!-- matplot(cbind(dat$ssb/1e9, res$summaries$SSB[1:200]/1e9), ylab="SSB (MT)", xlab="Year", col=c("cyan", "black"), lty=1, type=c("p","l"), pch=1) -->

<!-- validate_ssb(dat, pop$par$F_spf, pop$par$f_spf_before) -->
<!-- ``` -->

### Bioenergetic model

Now, let's simulate the bioenergetic model and compare the results with the model of Dankel et al.

```{r fig.height=9}
library(fisheRy)
params_file = here::here("params/cod_params.ini")

fish = new(Fish, params_file)
fish$par$s0 = 0.0192

pop = new(Stock, fish)
pop$readParams(params_file, F)
pop$superfish_size = 1e6  # Each superfish contains so many fish

v = pop$equilibriate_debug(5.61) |> matrix(nrow=200, byrow=T) |> as.data.frame() |> setNames(c("ssb", "tsb", "maturity", "nr", "dr", "nfish"))

d = pop$get_state()
dist = table(d$age, d$length)

par(mfrow = c(5,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5)
plot(v$nfish~seq(1,200,1), ylab="No. of superfish", xlab="Year", type="l", col=c("cyan3"))
plot(v$maturity~seq(1,200,1), ylab="Maturity", xlab="Year", type="l", col=c("cyan3"))
plot(I(v$nr/1e9)~seq(1,200,1), ylab="Recruitment", xlab="Year", type="l", col=c("cyan3"))
plot(v$dr~seq(1,200,1), ylab="actual/potential", xlab="Year", type="l", col=c("cyan3"))

res = simulate(0, 45, F)
matplot(cbind(v$ssb/1e9, res$summaries$SSB[1:200]/1e9), ylab="SSB (MT)", xlab="Year", col=c("cyan3", "black"), lty=1, type=c("p","l"), pch=1)

```

```{r fig.height=5}


par(mfrow = c(1,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5)
image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(1+dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length")



# validate_ssb(dat, pop$par$F_spf, pop$par$f_spf_before)
```


<!-- ### Bioenergetic model (with spawning mortality and spf) -->

<!-- Now, let's simulate the bioenergetic model and compare the results with the model of Dankel et al. -->

<!-- ```{r} -->
<!-- library(fisheRy) -->
<!-- params_file = "../params/cod_params.ini" -->

<!-- fish = new(Fish, params_file) -->
<!-- fish$par$s0 = 0.059 -->
<!-- fish$par$Mspawning = 1 -->

<!-- pop = new(Population, fish)   -->
<!-- pop$readParams(params_file, F) -->
<!-- pop$par$n = 1e6  # Each superfish contains so many fish -->
<!-- pop$init(1000, 1.93e3, 5.61)  # initialize the population with 1000 agents (superfish) -->
<!-- pop$par$F_spf = 1.0 -->

<!-- pop$verbose = T -->
<!-- nsteps = 200             # Let's simulate for 200 years -->
<!-- nfish = numeric(nsteps)  # Let's keep track of the number of superfish -->
<!-- nfish[1]=1000            # Since we initialized the population with 1000 superfish -->
<!-- cnames = pop$colnames -->
<!-- dat = data.frame(matrix(ncol=length(cnames), nrow=0)) -->
<!-- colnames(dat) = cnames -->
<!-- for (i in 1:nsteps){ -->
<!--   v = pop$update(5.61)       # Update all fish over 1 year  -->
<!--   dat[i,] = v    # some book-keeping -->
<!--   nfish[i] = pop$nfish() -->
<!-- } -->

<!-- d = pop$get_state() -->
<!-- dist = table(d$age, d$length) -->

<!-- par(mfrow = c(3,1), mar=c(5,5,1,1), oma=c(1,1,1,1), cex.lab=1.5, cex.axis=1.5) -->
<!-- plot(nfish~seq(1,nsteps,1), ylab="No. of superfish", xlab="Year") -->
<!-- image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length") -->

<!-- res = simulate(0, 45, F) -->
<!-- matplot(cbind(dat$ssb/1e9, res$summaries$SSB[1:200]/1e9), ylab="SSB (MT)", xlab="Year", col=c("cyan", "black"), lty=1, type=c("p","l"), pch=1) -->

<!-- validate_ssb(dat, pop$par$F_spf, pop$par$f_spf_before) -->
<!-- ``` -->



<!-- ### Quickly simulating an unfished population to equilibrium -->

<!-- It is possible to manually simulate a non-fished population to equilibrium as demonstrated in the previous tutorial. But there's an easier way: to quickly simulate an unfished population, we can use the function `noFishingEquilibriate()`. -->

<!-- ```{r} -->
<!-- # Create a population -->
<!-- pop_K = new(Population, fish) -->
<!-- pop_K$set_superFishSize(20e6) -->

<!-- # Equilibriate under no-fishing conditions -->
<!-- pop_K$verbose = T -->
<!-- state_equil = pop_K$noFishingEquilibriate(1.93e3, 5.61) -->

<!-- d = pop_K$get_state() -->
<!-- dist = table(d$age, d$length) -->

<!-- image(x=as.numeric(rownames(dist)), y = as.numeric(colnames(dist)), z=log(1+3*log(dist)), col=scales::viridis_pal()(100), xlab="Age", ylab="Length") -->


<!-- ``` -->



